---
title: Visualizing Strava Activities with R
author: Eoin Larkin
date: '2020-10-01'
slug: strava-r
categories: []
tags: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Strava is an internet service for tracking human exercise which incorporates social network features - it is sometimes referred to as Facebook for cyclist. Strava allows the visualisation of sports activities

Process:
- Bulk export data from Strava using the 
- Covert exported data to gpx files with GPSBabel
- Import data into R and use the Strava library by Marcus Volz to visualise the data [Link](https://github.com/marcusvolz)


The `strava` library makes visualizing and creating heatmaps of activities super simple. Howeer, it does require the activitiy files to be in `gpx` file format. This can cause some issues as the Strava bulk export can be in multiple file formats including `.fit` and `.tcx` depending on the make and model of the device that was used to record the activitiy.

Conversion of the files to the `gpx` file format is relatively straightforward; however at the time of writing the Strava output format for `tcx` files is not fully consistent with the file format specification. This issues is detailed in more detail here: [Link](https://github.com/gpsbabel/gpsbabel/issues/371) 

GPSBabel is very well documented and has cross platform support; it can be downloaded from here: [Link](https://www.gpsbabel.org/) 

Details on using GPSBabel to batch covert a set of files in Windows can be found here :[Link](https://www.gpsbabel.org/os/Windows_batch.html)

```{r include=FALSE}
# Installing the strava package
# devtools::install_github("marcusvolz/strava")

# Setting the path for the activities file
path <- "C:\\Users\\Eoin\\OneDrive\\Data Science\\Projects\\2020.12 - Strava Visualisation\\strava_activities_converted\\"
```

The first step is to load gpx activity files into Strava. This can be achieved by using the `process_data()` function from the `strava` package. The function converts the activity files into a single data frame which can then be easily plotted using `ggplot`.

```{r preliminaries}
library(strava) # loading the Strava package
library(patchwork) # for aligning the plots
data <- process_data(path) # loading the activities
```

## Facets Plot

The facets plot is the first of the plots that can be easily created using the package. Each activity file is visualised individually with a trace outline of the route of each of the activities.

```{r }
p1 <- plot_facets(data) + 
  theme(plot.margin = margin(0.5,0.5,0.5,0.5, "cm"))

ggsave("facets001.png", p1, width = 10, height = 10, units = "cm", dpi=1200)

p1
```

## Heatmap Plot

Next, the `plot_map` function makes creating heatmaps super simple. As part of the plotting, option latitude and longitude coordinates can be defined. As the library uses standard `ggplot` functions, additional code can be included to add titles and to modify the plot margins.

```{r}

# Defining our list of locations
loc <- list(dublin=c(-6.86, -6.00, 53.03, 53.46), 
                  trieste=c(13.42, 14.28, 45.47, 45.9), 
                  galway=c(-8.56, -9.78, 52.85, 53.46))

# Trieste
p2 <- plot_map(data, lon_min = loc$trieste[1], lon_max = loc$trieste[2], lat_min = loc$trieste[3], lat_max = loc$trieste[4]) +
      theme(plot.margin = margin(0.25,0.25,0.25,0.25, "cm")) +
      theme(plot.title = element_text(color = "gray40", size = 16, face = "bold"),
                            plot.subtitle = element_text(color = "gray40", size=3)) +
      theme(panel.border = element_rect(colour = "gray40", fill=NA, size=1)) 
      #+ labs(title = "Trieste")
                           
# Dublin
p3 <- plot_map(data, lon_min = loc$dublin[1], lon_max = loc$dublin[2], lat_min = loc$dublin[3], lat_max = loc$dublin[4]) +
      theme(plot.margin = margin(0.25,0.25,0.25,0.25, "cm")) +
      theme(plot.title = element_text(color = "gray40", size = 16, face = "bold"),
                            plot.subtitle = element_text(color = "gray40", size=3)) +
      theme(panel.border = element_rect(colour = "gray40", fill=NA, size=1)) 
      # + labs(title = "Dublin") 

# Galway
p4 <- plot_map(data, lon_min = loc$galway[1], lon_max = loc$galway[2], lat_min = loc$galway[3], lat_max = loc$galway[4]) +
      theme(plot.margin = margin(0.25,0.25,0.25,0.25, "cm")) +
      theme(plot.title = element_text(color = "gray40", size = 16, face = "bold"),
                            plot.subtitle = element_text(color = "gray40", size=3)) +
      theme(panel.border = element_rect(colour = "gray40", fill=NA, size=1)) 
      # + labs(title = "Galway")
  
# Arranging the plots using the patchwork package
plot_merge <- (p2 / p3 / p4) # creates a merged stacked plot
plot_merge

# Saving the plot output
# ggsave("map_merge.png", plot_merge, width = 10, height = 25, units = "cm",)
ggsave("map_merge.png", plot_merge)
ggsave("map_trieste.png", p2) # individually saving the plot
ggsave("map_dublin.png", p3)  # individually saving the plot 
ggsave("map_galway.png", p4) # individually saving the plot
```

## Activity Timeline

Activities can also be visualised in a calendar format:

```{r timeline_plot}
p5 <- plot_calendar(data, unit = "distance")

# Saving the plot output
ggsave("calendar001.png", p4, width = 7, height = 7, units = "cm")
```

## Ridges Plot

The ridges plot visualises activities by time across a seven day week:

```{r ridges_plot}
p6 <- plot_ridges(data)
ggsave("ridges001.png", p5, width = 20, height = 20, units = "cm")
```

## Packed Circles Plot

Finally the packed circles plot visualises activities by distance and year:

```{r packed_circle_plot}
p7 <- plot_packed_circles(data)
ggsave("packed_circles001.png", p6, width = 20, height = 20, units = "cm")
```

